---
title: "Regression Model Project"
author: "Expedito Pinto de Paula Junior"
date: "10/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Summary

Motor Trend Magazine about the automobile industry are looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

1. **"Is an automatic or manual transmission better for MPG?"**
2. **"Quantify the MPG difference between automatic and manual transmissions?"**

# 2. Exploratory Data Analysis

## 2.1 Load Libraries and Data Set
```{r load}
library(datasets)
library(ggplot2)
data("mtcars")
```

## 2.2 Basic Exploratory Data Analysis
```{r basics}
# Sample Data
head(mtcars)

# Variables
str(mtcars)

# summary
summary(mtcars)
```

## 2.3 Automatic x Manual Transmission

```{r automaticXmanual}
mtcars$am <- as.factor(mtcars$am)

ggplot(aes(x=am, y=mpg), data=mtcars) + 
     geom_boxplot(aes(fill=am)) + 
     labs(title = "Automatic x Manual Transmission") + 
     xlab("Transmission") + 
     ylab("mpg") + 
     labs(fill="0=Automatic, 1=Manual")
```

The plot above indicates that manual transmission provide better mileage per galon than the automatics. 

```{r t-Test}
AutomaticT <- mtcars[mtcars$am == "0",]
ManualT <- mtcars[mtcars$am == "1",]
t.test(AutomaticT$mpg, ManualT$mpg)
```

The hypothesis test performed rejects the null hypothesis that the transmission type is in fact significantly correlateed to mileage per gallon (mpg).

# 3.Regression Models

## 3.1 Linear Regression Model
```{r linearRM}
lmModel <- lm(mpg ~ am, data=mtcars)
summary(lmModel)
```

The linear regression shows how much transmission type affect mpg performance. Specifically manual transmissions provide 7.25 MPG better performance than automatics. But, based in R-squares value, transmission types only explain 36% of the MPG performance, based on this a simple linear regression is not enought to answer the Motor Trend's questions.

## 3.2 Multivariable Regression Model
```{r mVariable}
mModel <- lm(mpg ~ am + cyl + hp + wt, data = mtcars)
anova(lmModel, mModel)
summary(mModel)
```

The multivariable regression model (MRM) includes other factors (variables) that most likely affect a vehicle's including number of cylinders, engine horsepower, vehicle weight and others. The MRM shows that:

1. Each 1.48 MPG increase from manual transmission over automatic.
2. The multivariable model explain 85% of the MPG performance.

# 4. Conclusions

1. **"Is an automatic or manual transmission better for MPG?"**
A: Based in MPG, manual transmission perform better than automatic.

2. **"Quantify the MPG difference between automatic and manual transmissions?"**
A: In the first model, manual transmission perform better than automatic by 7.25 MPG, however this factor explain only 36% of the relation. When measure MPG adding cylinders, horsepower and weight variable, the manual transmission provide an additional 1.48 MPG of performance over automatic accounting 85% of the relation explanation. 

# 5. Appendix

As appendix is showed a pairs plot based in mpg and residual from multivariable regression model.

```{r PairsAndResidual}
pairs(mpg ~ ., data = mtcars)
par(mfrow = c(2,2))
plot(mModel)
```